Learning Auditory Similarities Between Instrumental Playing Techniques

### 2019-10-09 paper/paper.tex
details generation of sim matrices, and organizing figures

% \ml{vl citer, dire que c'est du coup une base solide de représentation et que le modèle mathamtique que l'on utilisera ici est quasi equivalent}.

%\ml{parler des travaux de Mounia}

@@

To construct a dataset for exploring the perception of timbre, we conducted a survey on the perceived similarity of a set of sounds.
The set consists of $78$ recordings of different instrument--PT (IPT) pairs, each played at the same pitch (C4) and nuance \textit{mezzo-forte}. For more information on the dataset considered in this study, see Section \ref{sec:dataset}.
In the survey, the subject is asked to organize dots on a two-dimensional place in a free sorting task.
Each dot corresponded to a different IPT recording and is moved by the user dragging them with the mouse.
In addition to spatially arranging the dots, the user is also asked to assign one of $20$ colors to each dot, providing a simple information of clustering of the sounds.
While both the location and color of each dot was recorded, only the color is used in our subsequent analysis---the spatial arrangement is only used to facilitate the user's ability to organize the sounds.
More information on the data collection process is available in  Section \ref{sec:judgments}.

The above survey was implemented as an interactive website, links to which were sent out on internal mailing lists of the \emph{Conservatoire national sup\'erieur de musique et de danse de Paris} (CNSMDP) and international mailing lists focusing on research in audio and music processing.
We made the experiment available for two months, during which we obtained results for a total of $31$ subjects.
The final arrangement and color assignment of one of the subjects is shown in Figure \ref{fix:xp2display}.

@@

It may also be of interest to condition the space using the aggregate similarity judgments given by several subjects. We found that using a hypergraph partitioning algorithm \cite{kernighan1970efficient,han1997scalable,strehl2002cluster} to produce the consensus similarity matrix performed well. Interestingly, although several sounds pairs were obtained from the same instrument (but different PTs) and others were obtained using the same PT (but different instruments), the consensus clustering did not favor one of these taxonomies over the other, see Figure \ref{fig:consensusVsIpt}.

\begin{figure}
\center
\includegraphics[width = \textwidth]{../code/xp2/consensusVsI.png}
\caption{'ASax'    'BBTb'    'BbCl'    'Bn'    'Cb'    'Fl'    'Gtr'    'Hn'    'Hp'    'Ob'    'TTbn'    'TpC'  'Va'    'Vc'}
\label{fig:consensusVsIpt}
\end{figure}

\begin{figure}
\center
\includegraphics[width = \textwidth]{../code/xp2/consensusVsPt.png}
\caption{ 'art-harm'    'blow-no-reed'    'explo-slap'    'harm-fngr'    'key-cl'    'legno-batt'    'legno-tratto'
    'nonvib'    'ord'    'ord-closed'    'ord-hi-reg'    'ord-open'    'pdl-tone'    'pizz'    'pizz-bartok'
    'pizz-lv'    'pizz-sec'    'pont'    'slap'    'slap-unp'    'tasto'    'tng-ram'}
\label{fig:consensusVsIpt}
\end{figure}

\ja{We should have some figure to back this up.
Can this be done using the tag clouds?}
\ml{tentative solution using similarity matrices, see fig, if ok should be improved by vl (code/xp2/consensusVsIpt.m)}

@@

@@

The second block $B_2$ considers one vector of features and the consensus clustering $D_c$ as inputs and another vector of features as output. For this block, the main factor is the type of projection $B_2^p \in \{ \text{lda}, \text{lmnn} \}$. The lda is the linear discriminant analysis \cite{} and the lmnn is the large margin nearest neighbors algorithm presented in Section \ref{}. For this algorithm the neighborhood is set to $5$.

@@

\section*{Figures}

\begin{figure}[h!]
\caption{\csentence{Model overview}
One line music players (3 ipts bass pizz, bass archo, violin archo) sound, spec,
scat, isomap.
Second line : perceptual clustering.}
\end{figure}

\begin{figure}[h!]
\caption{\csentence{Tag clouds}
A short description of the figure content should go here.}
\end{figure}

\begin{figure}[h!]
\caption{\csentence{Perfs MFCC, separable, joint (last)}
A short description of the figure content should go here.}
\end{figure}

@@

\section*{Tables}
\begin{table}[h!]
\caption{Sample table title. This is where the description of the table should go.}
      \begin{tabular}{cccc}
        \hline
           & B1  &B2   & B3\\ \hline
        A1 & 0.1 & 0.2 & 0.3\\
        A2 & ... & ..  & .\\
        A3 & ..  & .   & .\\ \hline
      \end{tabular}
\end{table}

>>>>

\ml{vl on fait toujours cette figure ?}

@@

To construct a dataset for exploring the perception of timbre, we conducted a survey on the perceived similarity of a set of sounds.
The set consists of $78$ recordings of different instrument--PT (IPT) pairs, each played at the same pitch (C4) and nuance \textit{mezzo-forte}. For more information on the dataset considered in this study, see Section \ref{sec:dataset}.
In the survey, the subject is asked to organize dots on a two-dimensional place in a free sorting task.
Each dot corresponded to a different IPT recording and is moved by the user dragging them with the mouse.
In addition to spatially arranging the dots, the user is also asked to assign one of $20$ colors to each dot, providing a simple information of clustering of the sounds.
While both the location and color of each dot was recorded, only the color is used in our subsequent analysis---the spatial arrangement is only used to facilitate the user's ability to organize the sounds.
More information on the data collection process is available in  Section \ref{sec:subjective}.

The above survey was implemented as an interactive website, links to which were sent out on internal mailing lists of the \emph{Conservatoire national sup\'erieur de musique et de danse de Paris} (CNSMDP) and international mailing lists focusing on research in audio and music processing.
We made the experiment available for two months, during which we obtained results for a total of $31$ subjects.
The final arrangement and color assignment of one of the subjects is shown in Figure \ref{fig:xp2display}.

@@

It may also be of interest to condition the space using the aggregate similarity judgments given by several subjects. We found that using a hypergraph partitioning algorithm \cite{kernighan1970efficient,han1997scalable,strehl2002cluster} to produce the consensus similarity matrix performed well.

In order to better interpret the information captured by the consensus clustering, we generate two similarity matrices, one for the 15 instruments used, see Figure \ref{fig:consensusVsI}, and one for the 23 PTs used, see Figure \ref{fig:consensusVsPt}. For the instruments similarity matrix, the value within each cell counts the number of times two instruments have been grouped in the consensus clustering, normalized by the average number of IPTs that considered the two instruments. The PTs similarity matrix is computed in the same way.

The ordering of the instruments or the PTs on the vertical and horizontals axes is chosen according to the ordering given by an agglomerative hierarchical clustering using the ward criterion of the similarity matrix under inspection. This ordering allows us to interpret the similarity matrices more easily compared to an alphabetical ordering while preserving the  similarity values between each pair of instruments or PTs.

Interestingly, although several sounds pairs were obtained from the same instrument (but different PTs) and others were obtained using the same PT (but different instruments), the consensus clustering did not favor one of these taxonomies over the other.

\ml{vl : je pense qu'il faut remplacer la derniere phrase par quelque chose de plus musicologique, en disant que l'on retrouve bien l'organologie, et montrer certains points d'interet sur les pts pour guider le lecteur.}

\begin{figure}
\center
\includegraphics[width = \textwidth]{figures/consensusVsI.png}
\caption{}
\label{fig:consensusVsI}
\end{figure}

\begin{figure}
\center
\includegraphics[width = \textwidth]{figures/consensusVsPt.png}
\caption{ }
\label{fig:consensusVsPt}
\end{figure}

@@

\begin{figure}[h!]
\caption{\csentence{Model overview}
One line music players (3 ipts bass pizz, bass archo, violin archo) sound, spec,
scat, isomap.
Second line : perceptual clustering.}
\label{fig:pipeline}
\end{figure}

@@

The second block $B_2$ considers one vector of features and the consensus clustering $D_c$ as inputs and another vector of features as output. For this block, the main factor is the type of projection $B_2^p \in \{ \text{lda}, \text{lmnn} \}$. The lda is the linear discriminant analysis \cite{} and the lmnn is the large margin nearest neighbors algorithm presented in Section \ref{sec:weighting}. For this algorithm the neighborhood is set to $5$.

@@

%\section*{Figures}

@@

% \section*{Tables}
% \begin{table}[h!]
% \caption{Sample table title. This is where the description of the table should go.}
%       \begin{tabular}{cccc}
%         \hline
%            & B1  &B2   & B3\\ \hline
%         A1 & 0.1 & 0.2 & 0.3\\
%         A2 & ... & ..  & .\\
%         A3 & ..  & .   & .\\ \hline
%       \end{tabular}
% \end{table}
